Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-16 demystify.tex
updates

\subsubsection{Types of Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.  The main difference is their focus: research, teaching, or both. Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students. Thus when contacting a prof. for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).

\textbf{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.  They often have a low teaching load, e.g., 1 course per semester. \emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).  \emph{Tenure-track} faculty are professors who are on the track to get tenure.  For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students. \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.

\textbf{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching. They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants. As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.

@@

Some universities have \textbf{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
They can recruit students and provide RA funding for them. 
They can also teach classes but typically have a lower teaching load.
For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.

\textbf{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun. \textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.   Due to their roles, these profs. typically do not advise PhD students.

\subsubsection{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.  Assistant means new faculty, associate means they have been promoted (or was hired after being in the industry for a while), and full means they are senior.

Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years). Note that getting tenure is a huge deal, and it might change the way they work with students, e.g., they might be more relaxed and less stressed (\autoref{sec:tenure-vs-tenure-track}).

>>>>

\paragraph{Faculty} There are typically two types of faculty at an R1 research university: tenure-line and teaching.  The main difference is their focus: research, teaching, or both. Tenure-line faculty are typically more research-focused and can mentor PhD students, while teaching faculty are more teaching-focused and typically do not advise PhD students. Thus when contacting a prof. for research opportunities (\autoref{sec:contact}), you should focus on tenure-line faculty (and not adjunct or emeritus :p).

\emph{Tenure-line faculty}, consisting of tenured and tenure-track profs., focuses on research, which includes publishing papers, obtaining grants, and mentoring PhD students.  They often have a low teaching load, e.g., 1 course per semester. \emph{Tenured} faculty are professors who have been promoted to permanent positions (informally, very hard to fire them).  \emph{Tenure-track} faculty are professors who are on the track to get tenure.  For most schools, tenure-line is the main type of faculty that can serve as a formal adviser of PhD students. \autoref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your adviser.

\emph{Teaching faculty}, sometimes called instructional faculty or professor of practice, mainly focuses on teaching. They typically teach 3--4 classes per semester (which is quite a lot) and \emph{do not} have research responsibilities, e.g., they do not have to worry much about publishing papers or obtaining grants. As such they typically do not have funding (e.g., RA~\autoref{sec:funding}) to support PhD students.
Depending on the university, they might be able to \emph{co-advise} and mentor PhD students with a tenure-line faculty, but they typically do not formally advise and graduate PhD students.

@@

Some universities have \emph{research faculty} or research scientists, who are typically non-tenure line faculty and focus on research (e.g., publishing papers and getting grants). 
They can recruit students and provide RA funding for them. 
They can also teach classes but typically have a lower teaching load.
For some universities, they \emph{can} solely advise and graduate PhD students, but for others, they cannot.

\emph{Adjunct faculty} is not full-time, e.g., they might be working in industry and teach a class or two for fun. \textbf{Emeritus faculty} means they are retired from the university, e.g., they could move to the industry, but still have some affiliation with the university.   Due to their roles, these profs. typically do not advise PhD students.

\paragraph{Ranks} Faculty have ranks such as assistant, associate, and full, regardless if they are tenure-line or teaching.  Assistant means new faculty, associate means they have been promoted (or was hired after being in the industry for a while), and full means they are senior.

Typically, a tenure-track faculty starts with being an assistant professor and then gets tenure and promoted to associate (around 6 years). Note that getting tenure is a huge deal, and it might change the way they work with students (\autoref{sec:tenure-vs-tenure-track}).
